### 2020-06-04 report/sections/server/node/runtime.tex
Report: progress on chapter 4 receive edge case

Ideally, a more succinct (and direct) representation would be
\begin{lstlisting}[language=javascript]
const successor = this.handler[decoded.label](...decoded.payload);
return next(successor);
\end{lstlisting}

But this expresses a type dependency between \texttt{label}
and \texttt{payload}, which as discussed, cannot be implemented.
But message structures \textit{precisely} define a discriminated
union, so we handle this with a switch statement,
at the cost of having the same code in each case -- but
TypeScript does infer different types in each case statement,
so the duplicated code does serve a functional purpose.

Returning to \cref{lst:noderuntimewsmsg},
note that the type of the \texttt{handler} property is \textit{optional}
-- this hints at a consistency problem: 
\textit{how do we know that \texttt{this.handler} is set when
a message is received?}

>>>>

Ideally, a more succinct (and direct) representation would be
\begin{lstlisting}[language=javascript,tabsize=2]
(message: any) => {
	const decoded = JSON.parse(message) as Message.S51;
	const successor =  this.handler[decoded.label](...decoded.payload);
	return next(successor);
}
\end{lstlisting}

But this expresses a type dependency between \texttt{label}
and \texttt{payload}, which as discussed 
(\cref{subsubsection:dependenttypes}), cannot be implemented.
However message structures \textit{precisely} 
define a discriminated union 
(the label acts as the discriminant 
to distinguish between payload types),
so we handle this with a switch statement,
at the cost of having the same code in each case --
TypeScript does infer the correct specific type in each case statement,
so code duplication here does serve a functional purpose.

Returning to \cref{lst:noderuntimewsmsg},
note that the type of the \texttt{handler} property is \textit{optional}
-- this hints at a problem: 
\textit{how do we know that \texttt{this.handler} is set when
a message is received?} The types imply that we \textit{do not},
and this is indeed the case.
In fact, when we consider a \textit{multiparty} context,
our approach with receive handler registration 
using an optional value actually \textit{fails} to guarantee correctness.
We motivate the problem with a worked
example.

\begin{example}[``Out-of-order'' message receives]
Recall that Node.js is a single-threaded event loop runtime, so
when a message arrives, 
the \texttt{onmessage} event is \textit{queued},
and current execution is \textit{not pre-empted}.

Now consider a multiparty session specified by the global type

\[ A \to S: \tmsg{M1(string)}. B \to S: \tmsg{M2(number)}. \texttt{end} \]

Given that \trole{S} is the server endpoint,
we describe a possible execution flow for the protocol
that breaks our implementation:

\begin{enumerate}
\item 
\trole{S} transitions to its initial state, ``receive from \trole{A}''.
The receive handler for \tmsg{M1} is registered.

\item
\tmsg{M2} arrives at \trole{S}, so the \texttt{onmessage} handler
is queued. This is \textit{perfectly plausible}: there
is no causal relation between \tmsg{M1} and \tmsg{M2}.

\item
\tmsg{M1} arrives at \trole{S}, so the \texttt{onmessage} handler
is queued.

\item
The \texttt{onmessage} handler for \tmsg{M2} is executed.
The registered handler expects \tmsg{M1(string)},
but it is callled with \tmsg{M2(number)}, which raises
a runtime type error.
\end{enumerate}

This exposes two problems: 
\textbf{(1)} the 
\end{example}
